---
title: "Modeling with Regression"
author: "Kobbe Harwell"
date: ""
output:
    html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r packages, warning = FALSE, message = FALSE}
library(ggplot2)
library(PASWR2)
library(car)
library(dplyr) 
library(lattice)
library(boot)
library(MASS)
```



## Data and idea comes from (*Militino et al., 2004*).


The goal is to create a multiple linear regression model that can be used to predict the total price of
apartments by their structural characteristics. The data frame `VIT2005` contains
several variables, and further description of the data can be found below.

A data frame with 218 observations on the following 5 variables:

* `totalprice` (the market total price (in Euros) of the apartment including garage(s) and storage room(s))

* `area` (the total living area of the apartment in square meters)

* `zone` (a factor indicating the neighborhood where the apartment is located with levels `Z11, Z21, Z31, Z32, Z34, Z35, Z36, Z37, Z38, Z41, Z42, Z43, Z44, Z45, Z46, Z47, Z48, Z49, Z52, Z53, Z56, Z61, and Z62`)

* `category` (a factor indicating the condition of the apartment with levels `2A, 2B, 3A, 3B, 4A, 4B`, and `5A` ordered so that `2A` is the best and `5A` is the worst)

* `age` (age of the apartment in years)

* `floor` (floor on which the apartment is located)

* `rooms` (total number of rooms including bedrooms, dining room, and kitchen)

* `out` (a factor indicating the percent of the apartment exposed to the elements: The levels `E100, E75, E50, and E25`, correspond to complete exposure, `75`% exposure, `50`% exposure, and `25`% exposure, respectively.)

* `conservation` (is an ordered factor indicating the state of conservation of the apartment. The levels `1A, 2A, 2B, and 3A` are ordered from best to worst conservation.)

* `toilets` (the number of bathrooms)

* `garage` (the number of garages)

* `elevator` (indicates the absence (0) or presence (1) of elevators.)

* `streetcategory` (an ordered factor from best to worst indicating the category of the street with levels `S2, S3, S4, and S5`)

* `heating` (a factor indicating the type of heating with levels `1A, 3A, 3B, and 4A` which correspond to: no heating, low-standard private heating, high-standard private heating, and central heating, respectively.)

* `storage` (the number of storage rooms outside of the apartment)


###

(a) Characterize the shape, center, and spread of the variable `totalprice`.

```{r}

ggplot(data = VIT2005, aes(x=totalprice)) + geom_density(fill="blue")+theme_bw()+geom_vline(aes(xintercept=median(totalprice)),color="red",linetype="dashed",size=1)

MD <- median(VIT2005$totalprice)
# 
iqr <- IQR(VIT2005$totalprice)
# 
c(MD, iqr)
```

**Observation:** The distribution of `totalprice` is skewed right with a median of `269750` and an `IQR` of `100125`


#########################################################################################################


(b) Explore the relationships between totalprice and the numerical explanatory variables `area, age, floor, rooms, toilets, garage, elevator`, and `storage`.

```{r}
scatterplotMatrix( ~ totalprice + area + age + floor + rooms, data = VIT2005)

scatterplotMatrix( ~ totalprice + toilets+ garage+ elevator+ storage, data = VIT2005)

```



(c) 

The correlation coefficients are:

```{r}
NUM <- c("area", "age", "floor", "rooms", "toilets", "garage","elevator", "storage")
COR <- cor(VIT2005[, "totalprice"], VIT2005[, NUM])
COR
```

**Observation:** The highest three correlations with `totalprice` occur with `area` (0.8092), `toilets` (0.6876), and `rooms`(0.5256).

**Model (A)** Backward elimination to develop a model that predicts totalprice using the data frame `VIT2005`. “P-value-to remove” of `5`%.

```{r}
model.be <- lm(totalprice ~ ., data = VIT2005)

drop1(model.be, test = "F")
```


Drop most insignificant variable from the model. 
```{r}
model.be <- update(model.be, .~. - age)
drop1(model.be, test = "F")
```

Which one is to be dropped next?

Answer: floor

```{r}
model.be <- update(model.be, .~. - floor)
drop1(model.be, test = "F")

```

Which one is to be dropped next?

Answer: conservation

```{r}
model.be <- update(model.be, .~. - conservation)
drop1(model.be, test = "F")

```

Which one is to be dropped next?

Answer: rooms

```{r}
model.be <- update(model.be, .~. - rooms)
drop1(model.be, test = "F")

```

Which one is to be dropped next?

Answer: storage

```{r}
model.be <- update(model.be, .~. - storage)
drop1(model.be, test = "F")
```


Finally,

```{r}
formula(model.be)

modelA <- lm(formula(model.be), data = VIT2005)
```


Observation: Backward elimination suggests using the variables `area`, `zone`, `category`, `out`, `toilets`, `garage`, `elevator`, `streetcategory`, and `heating` to best predict `totalprice`.



Compute $CV_n$, the leave-one-out cross validation error, for `modelA`. 
Compute $CV_5$, the **five-fold** cross validation error, for `modelA`. 
Compute $R^2$, $R^2_a$, the `AIC`, and the `BIC` for **Model (A)**.

(i)

```{r}
modelAg <- glm(formula(model.be), data = VIT2005)
cv.errorN <- cv.glm(VIT2005, modelAg)

CVNa <- cv.errorN$delta[1]
CVNa
```
 

```{r}
set.seed(5)
cv.error5 <- cv.glm(VIT2005, modelAg, K = 5)
CV5a <- cv.error5$delta[1]
CV5a
```

Observation: The $CV_n = 647026246$ for Model (A), and $CV_5 = 652546038$ for Model (A).


(ii)

```{r}

mgof <- function(model = model, data = DF, ...){
  R2a <- summary(model)$adj.r.squared
  R2 <- summary(model)$r.squared
  aic <- AIC(model)
  bic <- AIC(model, k = log(nrow(data)))
  se <- summary(model)$sigma
  form <- formula(model)
  ANS <- c(R2 = R2, R2.adj = R2a, AIC = aic, BIC = bic, SE = se)
  ANS
}

MGOF <- mgof(model = modelA, data = VIT2005)
MGOF
```

Observation: The total proportion of variability explained by modelA is `0.9138`.


(d)

Explore the residuals of Model A. (**Diagnostics: Checking the model assumptions**)

```{r}
residualPlot(modelA, main = "Model A")
```

Observation: The residuals versus the fitted values for Model (A) have a definite curvature indicating the model is not quite adequate.



**(e)** Finding a suitable transformation for totalprice and developing a new model

```{r}
boxCox(modelA, lambda = seq(-0.5, 0.5, length = 200))
```



#### Observation: A `log` transformation is suggested for the response `totalprice` in **Model (A)**.



#### **Model (E)**: Revised Log model. Data Frame `VIT2005`. “P-value-to remove” of `5`%.

```{r}
VIT2005$logtotalprice <- log(VIT2005$totalprice)

model.be <- lm(logtotalprice ~ ., data = VIT2005[ ,-1])
drop1(model.be, test = "F")

```

Decide to drop: age
```{r}
model.be <- update(model.be, .~. -age)
drop1(model.be, test = "F")
```

Decide to drop: conservation
```{r}
model.be <- update(model.be, .~. - conservation)
drop1(model.be, test = "F")
```

Decide to drop: floor
```{r}
model.be <- update(model.be, .~. - floor)
drop1(model.be, test = "F")
```


Decide to drop: rooms
```{r}
model.be <- update(model.be, .~. - rooms)
drop1(model.be, test = "F")
```

Decide to drop: streetcategory
```{r}
model.be <- update(model.be, .~. - streetcategory)
drop1(model.be, test = "F")
```



```{r}
formula(model.be)

modelE <- lm(formula(model.be), data = VIT2005)
```

```{r}
residualPlot(modelE, main = "Model E")
```